---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

#Importing Data
```{r}
lakedata=read.csv("RID_196_Chem_L375.csv")
summary(lakedata)
library(tidyverse)
View(lakedata)

# mutate a column that specifies before, during, and after aquaculture
before_data <- lakedata %>% 
  filter(year < 2003) %>% 
  mutate(Treatment = "Before")

during_data <- lakedata %>% 
  filter(year >= 2003 & year <= 2007) %>% 
  mutate(Treatment = "During")

after_data <- lakedata %>% 
  filter(year > 2007) %>% 
  mutate(Treatment = "After")

new_lakedata <- as.data.frame(rbind(before_data, during_data, after_data))
#View(new_lakedata)

# create a dataframe that has all observations before and during aquaculture
lakedata_no_recovery <- new_lakedata %>%
  filter(Treatment!="After") %>%
  filter(characteristic_name!="NA")
#View(lakedata_no_recovery)
```


#Creating a function to plot the average value of a specified variable in one year over time.
```{r}
generalplotfunction=function(z){
  data=new_lakedata%>%
    filter(characteristic_name==z)%>%
    group_by(year)%>%
    summarise(mean=mean(result_value))
graph=ggplot(data,aes(x=year,y=mean))+
    geom_line()+ geom_point()+geom_vline(xintercept = 2002,color = "blue", linewidth=1.5)+geom_vline(xintercept = 2008,  color = "blue", size=1.5)+labs(title=z)
return(graph)
}

```


#Plot all the variables 
```{r}
#target <- c(unique(lakedata$characteristic_name))
#significant_variables=filter(lakedata, characteristic_name %in% target)
#average_sigvariables_by_year=significant_variables%>%
#  group_by(year,characteristic_name)%>%
 # summarise(Mean=mean(result_value))
#ggplot(average_sigvariables_by_year,aes(x=year,y=Mean))+
#  geom_line()+ geom_point()+geom_vline(xintercept = 2002,color = "blue", size=1.5)+geom_vline(xintercept = 2008,  color = "blue", size=1.5)+facet_wrap(vars(characteristic_name),scales = "free_y")

characteristic.names=c(unique(lakedata$characteristic_name))
plot_list2=list()
for( i in 1:length(characteristic.names)){
   plot.data= generalplotfunction(characteristic.names[i])
 plot_list2[[i]] <- plot.data
}
plot_list2
```



#Plot the six main variables 
```{r}
library(ggpubr)

PARTP_timeseries <- generalplotfunction("PARTP")
PARTN_timeseries <- generalplotfunction("PARTN")
TDP_timeseries <- generalplotfunction("TDP")
SO4_timeseries <- generalplotfunction("SO4")
TDN_timeseries <- generalplotfunction("TDN")
CHLA_timeseries <- generalplotfunction("CHLA")

ggarrange(PARTP_timeseries, PARTN_timeseries, TDP_timeseries,
          SO4_timeseries, TDN_timeseries, CHLA_timeseries+ rremove("x.text"), 
          ncol = 2, nrow = 3)

# confused by this part of the code (Karen)
#unneeded can delete if nwe want.
target2=unique(new_lakedata$characteristic_name)
target2.vec=as.data.frame(target2)
 class(target2)
 
for (i in target2) {
  generalplotfunction(i)
}
```


#Plotting Histograms
```{r}
histogramfunction= function(charateristic)
{
  datafunc=new_lakedata%>%
    filter(characteristic_name==charateristic)%>%
    select(result_value)
  graph=ggplot(datafunc, aes(x = result_value)) +
    geom_histogram() +labs(title=charateristic)
  return(graph)
}

PARTP_hist <- histogramfunction("PARTP")
PARTN_hist <- histogramfunction("PARTN")
CHLA_hist <- histogramfunction("CHLA")
TDP_hist <- histogramfunction("TDP")
SO4_hist <- histogramfunction("SO4")
TDN_hist <- histogramfunction("TDN")

ggarrange(PARTP_hist, PARTN_hist, CHLA_hist, TDP_hist, SO4_hist,
          TDN_hist+ rremove("x.text"), 
          ncol = 2, nrow = 3)


# confused by this code here (Karen)
histogramfunction(target2[1])
for (i in target2) {
  histogramfunction(target2[i])
}

# creating a function that assesses the data's normality with shapiro test
normalfunction <- function(names) {
  n.data=new_lakedata %>%
    filter(characteristic_name==names) %>%
    select(result_value)
   num.data=as.numeric(n.data$result_value)
  norm.model=shapiro.test(num.data)
 return(norm.model$p.value)
}

# histograms for every variable 
characteristic.names=c(unique(new_lakedata$characteristic_name))
plot_list.norm <- list()
for( i in 1:length(characteristic.names)){
   plot.data <- histogramfunction(characteristic.names[i])
 plot_list.norm[[i]] <- plot.data
}
plot_list.norm
```

#Permutation
```{r}
#Inital Permutation code
permutation_function <- function(){
  reshuffled <- Partpdata
  reshuffled$TreatmentNEW <- sample(reshuffled$Treatment, size = nrow(reshuffled), replace = F)
  
  before <- reshuffled %>% subset(TreatmentNEW == "Before") %>% summarise(mean = mean(result_value))
  during <- reshuffled %>% subset(TreatmentNEW == "During") %>% summarise(mean = mean(result_value))

  return(before-during)
}

gaps <- c()

for (i in 1:1000){
  gaps[i] <- permutation_function()
}

hist(unlist(gaps))
```



#Creating function for LMM
```{r}
Chladata <- lakedata_no_recovery %>% 
  filter(characteristic_name == "CHLA")
SO4data <- lakedata_no_recovery %>%
  filter(characteristic_name == "SO4")

PARTP_lmm <- lmer(result_value~Treatment+(1|year), data= Partpdata, 
                  REML = FALSE)
TDP_lmm <- lmer(result_value~Treatment+(1|year), data= Tdpdata, 
                  REML = FALSE)
PARTN_lmm <- lmer(result_value~Treatment+(1|year), data= Partndata, 
                  REML = FALSE)
TDN_lmm <- lmer(result_value~Treatment+(1|year), data= TdNdata, 
                  REML = FALSE)
CHLA_lmm <- lmer(result_value~Treatment+(1|year), data= Chladata, 
                  REML = FALSE)
SO4_lmm <- lmer(result_value~Treatment+(1|year), data= SO4data, 
                  REML = FALSE)

# this isn't working right now
vectorofpvalues <- c()
for (i in 1:length(characteristic.names)){
  mixed_model_data <- lakedata_no_recovery %>% 
    filter(characteristic_name == characteristic.names[i]) %>% 
    select(result_value, Treatment, year)
  vectorofpvalues[i] <- summary(lmer(result_value~Treatment+(1|year), data=mixed_model_data, REML = FALSE))$coefficients[2,5]
}

#this works now, but double check implementation of year
mixedmodelfunction <- function(characteristic){
  mixed_model_data <- lakedata_no_recovery %>%
    filter(characteristic_name == characteristic) %>%
    select(result_value, Treatment, year)
 
 return(summary(lmer(result_value~Treatment+(1|year), data=mixed_model_data, REML=FALSE)))
}

mixedmodelfunction("PARTP")

# FOR PARTP (not useful, ignore)
models <- NULL

for (i in 1:length(unique(Partpdata$year))){
  data <- Partpdata %>% subset(year == i)
  models[[i]] <- cbind(
    summary(lmer(result_value~Treatment+(1|as.factor(Partpdata$year)), data = Partpdata))$coefficients[,c(1,4)],
    year = unique(Partpdata$year)[i])
}

do.call(rbind, models)

#Possible other methods
#this method returns the pvalue for a mixed model run for each characteristic except tss and na. The model takes in treatment before and after and the result value as the predictor and response variable. Year is taken as a randomn variable. We need to check exactly what the randomn variable thing means. I think we have a number of values in each year this thus treats the year a beach. Thus each year maybe independeant now. 
vectorofpvalues=c()
vectorofeffectsize=c()
characteristic.names=unique(lakedata_no_recovery$characteristic_name)
for (i in 1:(length(characteristic.names)-1)){
  mixed_model_data=lakedata_no_recovery%>%
    filter(characteristic_name==characteristic.names[i])%>%
    select(result_value,Treatment,year)
   vectorofpvalues[i]=(summary(lmer(result_value~Treatment+(1|as.factor(mixed_model_data$year)), data=mixed_model_data, REML=FALSE)))$coefficients[2,5]
 vectorofeffectsize[i]=(summary(lmer(result_value~Treatment+(1|as.factor(mixed_model_data$year)), data=mixed_model_data, REML=FALSE)))$coefficients[2,1]
}
pvalues_and_characteristic_name=data.frame(characteristic = characteristic.names[1:19], P.value = vectorofpvalues,signifcance.of.pvalue=signifcance,effect.size=vectorofeffectsize)
signifcance <- vectorofpvalues < 0.05
View(pvalues_and_characteristic_name)
```



#creating function for Permutation with effect size (DO NOT USE)
```{r}
library(tidyverse)
#Making Permutation Function
filterfunction=function(characteristic){
  data.perm=lakedata_no_recovery%>%
  filter(characteristic_name==characteristic)%>%
  select(Treatment,result_value,year)
  return(data.perm)
  }
permutation_function <- function(data12){
  reshuffled <- data12
  reshuffled$TreatmentNEW <- sample(reshuffled$Treatment, size = nrow(reshuffled), replace = F)
  model=glm(result_value~TreatmentNEW, data=reshuffled)
  summary.model=summary(model)
  return(summary.model$coefficients[2,1])
}


final.perm.function= function(characteristic){
filtered.data=filterfunction(characteristic)
estimates <- c()
for (i in 1:10000){
  estimates[i] <- permutation_function(filtered.data)
}
estimates=as.data.frame(estimates)
 data.perm=lakedata_no_recovery%>%
  filter(characteristic_name==characteristic)%>%
  select(Treatment,result_value,year)
model.new=summary(glm(result_value~Treatment, data = data.perm))
lol=model.new$coefficients[2,1]
plot_s=ggplot(estimates, aes(x=estimates)) + 
 geom_histogram( colour="black", fill="white")+ geom_vline(aes(xintercept=lol),
            color="blue", linetype="dashed", size=1)+labs(title =characteristic)
return(plot_s)
}
characteristic.names=c(unique(lakedata_no_recovery$characteristic_name))
plot_list <- list()
for( i in 1:(length(characteristic.names)-1)){
   plot.data= final.perm.function(characteristic.names[i])
 plot_list[[i]] <- plot.data
}

plot_list
```

#Permutation- difference in means between 'before' and 'during' treatments
```{r}
library(tidyverse)

filterfunction <- function(characteristic){
  data.perm=lakedata_no_recovery%>%
  filter(characteristic_name==characteristic)%>%
  select(Treatment,result_value,year)
  return(data.perm)
  }
permutation_function_mean <- function(data){
  reshuffled <- data
  reshuffled$TreatmentNEW <- sample(reshuffled$Treatment, size = nrow(reshuffled), replace = T)
    
  before <- reshuffled %>% subset(TreatmentNEW == "Before") %>% summarise(mean = mean(result_value))
  during <- reshuffled %>% subset(TreatmentNEW == "During") %>% summarise(mean = mean(result_value))
  bd=during-before
  return(bd[1,1])
}


final.perm.function <- function(characteristic){
filtered.data <- filterfunction(characteristic)
estimates <- c()
for (i in 1:1000){
  estimates[i] <- permutation_function_mean(filtered.data)
}
estimates <- as.data.frame(estimates)
 data.perm=lakedata_no_recovery%>%
  filter(characteristic_name==characteristic)%>%
  select(Treatment,result_value,year)
before.1 <- data.perm %>%  subset(Treatment == "Before") %>% summarise(mean = mean(result_value))
during.1 <-data.perm %>% subset(Treatment == "During") %>% summarise(mean = mean(result_value))
lol=(during.1-before.1)
lol[1,1]
plot_s=ggplot(estimates, aes(x=estimates)) + 
 geom_histogram( colour="black", fill="white")+
  labs(title=characteristic, x= "Difference in means", y= "Count") +
  geom_vline(aes(xintercept=lol[1,1]), color="blue", linetype="dashed", size=1) +
  geom_text(aes(x=(lol[1,1]-1.5), y=90, label = "Real difference"), angle = 90)
return(plot_s)
}
#characteristic.names=c(unique(lakedata_no_recovery$characteristic_name))
#plot_list <- list()
#for( i in 1:(length(characteristic.names)-1)){
#   plot.data= final.perm.function(characteristic.names[i])
# plot_list[[i]] <- plot.data
#}

plot_list

## Plotting for the two nitrogen variables we care about
characteristic.names

PARTN_perm_hist <- final.perm.function("PARTN"); PARTN_perm_hist
TDN_perm_hist <- final.perm.function("TDN"); TDN_perm_hist


## Plotting for the two phosphorus variables we care about

PARTP_perm_hist <- final.perm.function("PARTP"); PARTP_perm_hist
TDP_perm_hist <- final.perm.function("TDP"); TDP_perm_hist

## Plotting for CHLA
CHLA_perm_hist <- final.perm.function("CHLA"); CHLA_perm_hist




```


```{r}
lakedata_no_recovery %>% 
  filter(characteristic_name== "CHLA") %>%
  group_by(Treatment) %>% 
  summarize(means = mean(result_value))

```


#Function for ALL LMMs
```{r}
#Having issues with this code. The things run induvidually but not in the loop as a function# Ran this code in the chunk above. Works now. 
returnpvalueforllm=function(characteristic){
  datafortest=filterfunction(characteristic)
  model.new= lmer(result_value~Treatment+(1|as.factor(datafortest$year)), data=datafortest, REML=FALSE)
  summary.model=summary(model.new)
  return(summary.model)
}

returnpvalueforllm("ALK")
pvalues=c()
for( i in 1:(length(characteristic.names)-1)){
  pvalues[i]=  returnpvalueforllm(characteristic.names[i])

}
pvalues=c()

```


#Running Aarimas

```{r}
library(forecast)
no_recovery_forecastfunction=function(y){
  data1=lakedata_no_recovery %>%
    filter(characteristic_name==y)%>%
    select(result_value)
  model3=auto.arima(data1) 
  summary(model3)
  forecast_data3 <- forecast(model3, 20) 
  plot=plot(forecast_data3, main = y)
  return(plot)
}


beforeforecastfunction=function(y){
  data1=lakedata%>%
    filter(characteristic_name==y)%>%
    filter(year<2003)%>%
    select(result_value)
  model3=auto.arima(data1) 
  summary(model3)
  forecast_data3 <- forecast(model3, 20) 
  plot=plot(forecast_data3, main = y)
  return(plot)
}

duringforecastfunction=function(y){
  data1=lakedata%>%
    filter(characteristic_name==y)%>%
    filter(year<2007)%>%
    select(result_value)
  model3=auto.arima(data1) 
  summary(model3)
  forecast_data3 <- forecast(model3, 20) 
  plot=plot(forecast_data3, main = y)
  return(plot)
}


aarimalist <- list()
for( i in 1:(length(characteristic.names)-1)){
   plot.data= no_recovery_forecastfunction(characteristic.names[i])
 aarimalist[[i]] <- plot.data
}



beforeariimalist=list()
for( i in 1:(length(characteristic.names)-1)){
   plot.data= beforeforecastfunction(characteristic.names[i])
 beforeariimalist[[i]] <- plot.data
}

duringariimalist=list()
for( i in 1:(length(characteristic.names)-1)){
   plot.data= beforeforecastfunction(characteristic.names[i])
 duringariimalist[[i]] <- plot.data
}
ariimalist
```


#ALL DATA INTERPOLATED
```{r}
interpolationfunction=function(k){
  data3=lakedata%>%
    filter(characteristic_name== k)%>%
    group_by(year)%>%
    summarise(mean=mean(result_value))
  allyears=seq(1982,2022,1)
  dfallyears=data.frame(year=allyears)
  merged_data <- merge(dfallyears, data3, by = "year", all = TRUE)
  finaldata=merged_data%>%
    select(mean)
  na_p_data2 <- na.interp(finaldata)
  added_data_P=as.data.frame(na_p_data2)
  autoplot(na_p_data2, series="Interpolated") +
    autolayer(as.ts(finaldata), series="Original") +
    geom_point()+
    scale_colour_manual(values=c(`Interpolated`="red",`Original`="Blue"))+
    labs(title = k) +geom_vline(xintercept = c(22, 26), linetype = "dashed", color = "red")
}

library(forecast)
interplatedlist=list()
for( i in 1:(length(characteristic.names)-1)){
   plot.data= interpolationfunction(characteristic.names[i])
 interplatedlist[[i]] <- plot.data
}
interplatedlist
```

